Trends from 30-year observations of downward solar irradiance in Thessaloniki, Greece

### 2023-12-03 MDPI_submission.Rmd
Commit 2023-12-03 03:09

<!-- aggregation -->
In order to investigate the SDR trends, we implemented an appropriate aggregation scheme to the 1-minute data to derive a series in coarser time-scales.
To preserve the representativeness of the data we used the following criteria:
<!-- a) for daily mean values we exclude days with less than `r Daily_aggregation_N_lim` valid data points, -->
a) for the daily mean values we accept days with more than `r 100 * All_daily_ratio_lim`% of the daytime measurements present and valid,
b) monthly values were computed from daily means
only when at least `r Monthly_aggegation_N_lim` days were available.
To create the daily and monthly climatological means, we averaged the data based on the day of year and calendar month, respectively.
<!-- Seasonal -->
For the seasonal means we averaged the mean daily values in each season (Winter: December - February, Spring: March - May, etc.).
<!-- Deseasonalization -->
Finally, each data set was deseasonalized by subtracting the corresponding climatological annual cycle (daily or monthly) from the actual data.
To estimate the SZA effect on the SDR trends, the one-minute data were aggregated in $1^\circ$ SZA bins, separately for the morning and afternoon hours, and then were deseasonalized as mentioned above.

@@

## numeric precision of table
LT$slope    <- signif(LT$slope,   3)
LT$slope.p  <- signif(LT$slope.p, 3)
## rename sky conditions
LT$DATA <- sub(" cond.", "", LT$DATA)
LT$DATA <- sub(" sky", " skies", LT$DATA)

@@

\begin{figure}[h!]
    \begin{adjustwidth}{-\extralength}{0cm}
        {\centering 
            \subfloat[All skies.\label{fig:szatrends-1}]
                {\includegraphics[width=.32\linewidth]{./images/SzaTrends-1}}\hfill
            \subfloat[Clear skies.\label{fig:szatrends-2}]
                {\includegraphics[width=.32\linewidth]{./images/SzaTrends-4}}\hfill
            \subfloat[Cloudy skies.\label{fig:szatrends-3}]
                {\includegraphics[width=.32\linewidth]{./images/SzaTrends-7}}
        }
        \caption{Long term trends of SDR as a function of SZA separately form morning and afternoon periods. Solid shapes  represent statistically significant trends ($p < 0.005$).}\label{fig:szatrends}
    \end{adjustwidth}
\end{figure}

<!-- ```{r szatrends, echo = FALSE, fig.cap='Long term trends of SDR as a function of SZA separately form morning and afternoon periods. Solid shapes  represent statistically significant trends ($p < 0.005$).', fig.subcap=c('All skies.', 'Clear skies.', 'Cloudy skies.'), out.width='.32\\linewidth', fig.asp=1, fig.ncol = 3} -->
<!-- knitr::include_graphics("./images/SzaTrends-1.png") -->
<!-- knitr::include_graphics("./images/SzaTrends-4.png") -->
<!-- knitr::include_graphics("./images/SzaTrends-7.png") -->
<!-- ``` -->

@@

## numeric precision
spp$slope          <- signif(spp$slope,          3)
spp$slope.p        <- signif(spp$slope.p,        3)
spp$slope.stat_sig <- signif(spp$slope.stat_sig, 3)

>>>>

<!-- aggregation -->
In order to investigate the SDR trends, we implemented an appropriate aggregation scheme to the 1-minute data to derive a series in coarser time-scales.
To preserve the representativeness of the data we used the following criteria:
<!-- a) for daily mean values we exclude days with less than `r Daily_aggregation_N_lim` valid data points, -->
a) we accept only days with more than `r 100 * All_daily_ratio_lim`% of the daytime measurements present and valid,
b) on the dataset of clear- and cloudy-skies, we included only days with more than
`r 100 * Clear_daily_ratio_lim`% of the daytime measurements identified as clear or cloudy respectively,
c) monthly values were computed from daily means
only when at least `r Monthly_aggegation_N_lim` days were available.
To create the daily and monthly climatological means, we averaged the data based on the day of year and calendar month, respectively.
<!-- Seasonal -->
For the seasonal means we averaged the mean daily values in each season (Winter: December - February, Spring: March - May, etc.).
<!-- Deseasonalization -->
Finally, each data set was deseasonalized by subtracting the corresponding climatological annual cycle (daily or monthly) from the actual data.
To estimate the SZA effect on the SDR trends, the one-minute data were aggregated in $1^\circ$ SZA bins, separately for the morning and afternoon hours, and then were deseasonalized as mentioned above.

@@

## numeric precision of table
LT$slope    <- signif(LT$slope,   3)
LT$slope.p  <- signif(LT$slope.p, 3)
LT$Rsqrd    <- signif(LT$Rsqrd,   3)

LT$RsqrdAdj <- NULL
LT$cor.t    <- NULL

## rename sky conditions
LT$DATA <- sub(" cond.", "", LT$DATA)
LT$DATA <- sub(" sky", " skies", LT$DATA)

LT <- rename(LT, "Trend [%/year]" = slope)
LT <- rename(LT, "Sky condition"  = DATA )
LT <- rename(LT, "Days"           = N    )

@@

\begin{figure}[h!]
    \begin{adjustwidth}{-\extralength}{0cm}
        {\centering 
            \subfloat[All skies.\label{fig:szatrends-1}]
                {\includegraphics[width=.32\linewidth]{./images/SzaTrends-1}}\hfill
            \subfloat[Clear skies.\label{fig:szatrends-2}]
                {\includegraphics[width=.32\linewidth]{./images/SzaTrends-4}}\hfill
            \subfloat[Cloudy skies.\label{fig:szatrends-3}]
                {\includegraphics[width=.32\linewidth]{./images/SzaTrends-7}}
        }
        \caption{Long term trends of SDR as a function of SZA separately from morning and afternoon periods. Solid shapes  represent statistically significant trends ($p < 0.005$).}\label{fig:szatrends}
    \end{adjustwidth}
\end{figure}

<!-- ```{r szatrends, echo = FALSE, fig.cap='Long term trends of SDR as a function of SZA separately from morning and afternoon periods. Solid shapes  represent statistically significant trends ($p < 0.005$).', fig.subcap=c('All skies.', 'Clear skies.', 'Cloudy skies.'), out.width='.32\\linewidth', fig.asp=1, fig.ncol = 3} -->
<!-- knitr::include_graphics("./images/SzaTrends-1.png") -->
<!-- knitr::include_graphics("./images/SzaTrends-4.png") -->
<!-- knitr::include_graphics("./images/SzaTrends-7.png") -->
<!-- ``` -->

@@

## numeric precision
spp$slope          <- signif(spp$slope,          3)
spp$slope.p        <- signif(spp$slope.p,        3)
spp$slope.stat_sig <- signif(spp$slope.stat_sig, 3)
spp$Rsqrd          <- signif(spp$Rsqrd,          3)
spp$RsqrdAdj <- NULL
spp$cor.t    <- NULL

### 2023-12-03 MDPI_submission.tex
Commit 2023-12-03 03:09

In order to investigate the SDR trends, we implemented an appropriate
aggregation scheme to the 1-minute data to derive a series in coarser
time-scales. To preserve the representativeness of the data we used the
following criteria: a) for the daily mean values we accept days with
more than 50\% of the daytime measurements present and valid, b) monthly
values were computed from daily means only when at least 20 days were
available. To create the daily and monthly climatological means, we
averaged the data based on the day of year and calendar month,
respectively. For the seasonal means we averaged the mean daily values
in each season (Winter: December - February, Spring: March - May, etc.).
Finally, each data set was deseasonalized by subtracting the
corresponding climatological annual cycle (daily or monthly) from the
actual data. To estimate the SZA effect on the SDR trends, the
one-minute data were aggregated in \(1^\circ\) SZA bins, separately for
the morning and afternoon hours, and then were deseasonalized as
mentioned above.

@@

\caption{\label{tab:trendtable}Trends in SDR daily means for different sky conditions for the period 1993 - 2023.}
\begin{tabu} to \linewidth {>{\centering\arraybackslash}p{8em}>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X>{\centering}X>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X>{\centering}X}
\toprule
slope & slope.p & Rsqrd & RsqrdAdj & cor.t & cor.estimate & DATA & N & slope.stat\_sig\\
\midrule
0.3760 & 0.0000 & 0.008339 & 0.0082421 & 9.284 & 0.09132 & All skies & 10251 & 100.00\\
0.3430 & 0.0000 & 0.005312 & 0.0051756 & 6.234 & 0.07289 & Cloudy skies & 7279 & 100.00\\
0.0406 & 0.0141 & 0.001195 & 0.0009971 & 2.455 & 0.03458 & Clear skies & 5036 & 98.59\\
\bottomrule
\end{tabu}
\end{table}

@@

\begin{figure}[h!]
    \begin{adjustwidth}{-\extralength}{0cm}
        {\centering 
            \subfloat[All skies.\label{fig:szatrends-1}]
                {\includegraphics[width=.32\linewidth]{./images/SzaTrends-1}}\hfill
            \subfloat[Clear skies.\label{fig:szatrends-2}]
                {\includegraphics[width=.32\linewidth]{./images/SzaTrends-4}}\hfill
            \subfloat[Cloudy skies.\label{fig:szatrends-3}]
                {\includegraphics[width=.32\linewidth]{./images/SzaTrends-7}}
        }
        \caption{Long term trends of SDR as a function of SZA separately form morning and afternoon periods. Solid shapes  represent statistically significant trends ($p < 0.005$).}\label{fig:szatrends}
    \end{adjustwidth}
\end{figure}

@@

\caption{\label{tab:trendseasontable}SDR linear trends of monthly anomalies for each season of the year.}
\begin{tabu} to \linewidth {>{\centering\arraybackslash}p{8em}>{\centering}X>{\raggedleft}X>{\raggedleft}X>{\centering}X>{\centering}X>{\raggedleft}X>{\raggedleft}X>{\centering}X}
\toprule
Trend [\%/year] & slope.p & Rsqrd & RsqrdAdj & cor.t & cor.estimate & Sky condition & Season & slope.stat\_sig\\
\midrule
\cellcolor{gray!6}{0.6870} & \cellcolor{gray!6}{0.001330} & \cellcolor{gray!6}{0.3121839} & \cellcolor{gray!6}{0.2876190} & \cellcolor{gray!6}{3.564906} & \cellcolor{gray!6}{0.5587342} & \cellcolor{gray!6}{All skies} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{99.9}\\
\cmidrule{1-9}
0.1460 & 0.178000 & 0.0616942 & 0.0293388 & 1.380858 & 0.2483832 & All skies & Spring & 82.2\\
\cmidrule{1-9}
\cellcolor{gray!6}{0.1230} & \cellcolor{gray!6}{0.096100} & \cellcolor{gray!6}{0.0957525} & \cellcolor{gray!6}{0.0634580} & \cellcolor{gray!6}{1.721910} & \cellcolor{gray!6}{0.3094390} & \cellcolor{gray!6}{All skies} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{90.4}\\
\cmidrule{1-9}
0.4300 & 0.006670 & 0.2347063 & 0.2073744 & 2.930403 & 0.4844650 & All skies & Autumn & 99.3\\
\cmidrule{1-9}
\cellcolor{gray!6}{0.4990} & \cellcolor{gray!6}{0.000166} & \cellcolor{gray!6}{0.4025975} & \cellcolor{gray!6}{0.3812617} & \cellcolor{gray!6}{4.343912} & \cellcolor{gray!6}{0.6345057} & \cellcolor{gray!6}{Clear skies} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{100.0}\\
\cmidrule{1-9}
-0.0865 & 0.315000 & 0.0348099 & 0.0015275 & -1.022690 & -0.1865741 & Clear skies & Spring & 68.5\\
\cmidrule{1-9}
\cellcolor{gray!6}{-0.2180} & \cellcolor{gray!6}{0.000398} & \cellcolor{gray!6}{0.3659586} & \cellcolor{gray!6}{0.3433142} & \cellcolor{gray!6}{-4.020094} & \cellcolor{gray!6}{-0.6049451} & \cellcolor{gray!6}{Clear skies} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{100.0}\\
\cmidrule{1-9}
0.1090 & 0.221000 & 0.0529350 & 0.0191112 & 1.251009 & 0.2300761 & Clear skies & Autumn & 77.9\\
\cmidrule{1-9}
\cellcolor{gray!6}{0.5200} & \cellcolor{gray!6}{0.009580} & \cellcolor{gray!6}{0.2164332} & \cellcolor{gray!6}{0.1884487} & \cellcolor{gray!6}{2.781012} & \cellcolor{gray!6}{0.4652238} & \cellcolor{gray!6}{Cloudy skies} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{99.0}\\
\cmidrule{1-9}
0.1780 & 0.122000 & 0.0806332 & 0.0489309 & 1.594819 & 0.2839598 & Cloudy skies & Spring & 87.8\\
\cmidrule{1-9}
\cellcolor{gray!6}{0.1940} & \cellcolor{gray!6}{0.036800} & \cellcolor{gray!6}{0.1466137} & \cellcolor{gray!6}{0.1161356} & \cellcolor{gray!6}{2.193277} & \cellcolor{gray!6}{0.3829017} & \cellcolor{gray!6}{Cloudy skies} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{96.3}\\
\cmidrule{1-9}
0.3470 & 0.100000 & 0.0934930 & 0.0611177 & 1.699350 & 0.3057662 & Cloudy skies & Autumn & 90.0\\
\bottomrule
\end{tabu}
\end{table}

>>>>

In order to investigate the SDR trends, we implemented an appropriate
aggregation scheme to the 1-minute data to derive a series in coarser
time-scales. To preserve the representativeness of the data we used the
following criteria: a) we accept only days with more than 50\% of the
daytime measurements present and valid, b) on the dataset of clear- and
cloudy-skies, we included only days with more than 30\% of the daytime
measurements identified as clear or cloudy respectively, c) monthly
values were computed from daily means only when at least 20 days were
available. To create the daily and monthly climatological means, we
averaged the data based on the day of year and calendar month,
respectively. For the seasonal means we averaged the mean daily values
in each season (Winter: December - February, Spring: March - May, etc.).
Finally, each data set was deseasonalized by subtracting the
corresponding climatological annual cycle (daily or monthly) from the
actual data. To estimate the SZA effect on the SDR trends, the
one-minute data were aggregated in \(1^\circ\) SZA bins, separately for
the morning and afternoon hours, and then were deseasonalized as
mentioned above.

@@

\caption{\label{tab:trendtable}Trends in SDR daily means for different sky conditions for the period 1993 - 2023.}
\begin{tabu} to \linewidth {>{\centering\arraybackslash}p{8em}>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X>{\centering}X>{\raggedleft}X>{\raggedleft}X}
\toprule
Trend [\%/year] & slope.p & Rsqrd & cor.estimate & Sky condition & Days & slope.stat\_sig\\
\midrule
0.3760 & 0.0000 & 0.00834 & 0.09132 & All skies & 10251 & 100.00\\
0.3430 & 0.0000 & 0.00531 & 0.07289 & Cloudy skies & 7279 & 100.00\\
0.0406 & 0.0141 & 0.00120 & 0.03458 & Clear skies & 5036 & 98.59\\
\bottomrule
\end{tabu}
\end{table}

@@

\begin{figure}[h!]
    \begin{adjustwidth}{-\extralength}{0cm}
        {\centering 
            \subfloat[All skies.\label{fig:szatrends-1}]
                {\includegraphics[width=.32\linewidth]{./images/SzaTrends-1}}\hfill
            \subfloat[Clear skies.\label{fig:szatrends-2}]
                {\includegraphics[width=.32\linewidth]{./images/SzaTrends-4}}\hfill
            \subfloat[Cloudy skies.\label{fig:szatrends-3}]
                {\includegraphics[width=.32\linewidth]{./images/SzaTrends-7}}
        }
        \caption{Long term trends of SDR as a function of SZA separately from morning and afternoon periods. Solid shapes  represent statistically significant trends ($p < 0.005$).}\label{fig:szatrends}
    \end{adjustwidth}
\end{figure}

@@

\caption{\label{tab:trendseasontable}SDR linear trends of monthly anomalies for each season of the year.}
\begin{tabu} to \linewidth {>{\centering\arraybackslash}p{8em}>{\centering}X>{\raggedleft}X>{\raggedleft}X>{\centering}X>{\centering}X>{\raggedleft}X}
\toprule
Trend [\%/year] & slope.p & Rsqrd & cor.estimate & Sky condition & Season & slope.stat\_sig\\
\midrule
\cellcolor{gray!6}{0.6870} & \cellcolor{gray!6}{0.001330} & \cellcolor{gray!6}{0.3120} & \cellcolor{gray!6}{0.5587342} & \cellcolor{gray!6}{All skies} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{99.9}\\
\cmidrule{1-7}
0.1460 & 0.178000 & 0.0617 & 0.2483832 & All skies & Spring & 82.2\\
\cmidrule{1-7}
\cellcolor{gray!6}{0.1230} & \cellcolor{gray!6}{0.096100} & \cellcolor{gray!6}{0.0958} & \cellcolor{gray!6}{0.3094390} & \cellcolor{gray!6}{All skies} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{90.4}\\
\cmidrule{1-7}
0.4300 & 0.006670 & 0.2350 & 0.4844650 & All skies & Autumn & 99.3\\
\cmidrule{1-7}
\cellcolor{gray!6}{0.4990} & \cellcolor{gray!6}{0.000166} & \cellcolor{gray!6}{0.4030} & \cellcolor{gray!6}{0.6345057} & \cellcolor{gray!6}{Clear skies} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{100.0}\\
\cmidrule{1-7}
-0.0865 & 0.315000 & 0.0348 & -0.1865741 & Clear skies & Spring & 68.5\\
\cmidrule{1-7}
\cellcolor{gray!6}{-0.2180} & \cellcolor{gray!6}{0.000398} & \cellcolor{gray!6}{0.3660} & \cellcolor{gray!6}{-0.6049451} & \cellcolor{gray!6}{Clear skies} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{100.0}\\
\cmidrule{1-7}
0.1090 & 0.221000 & 0.0529 & 0.2300761 & Clear skies & Autumn & 77.9\\
\cmidrule{1-7}
\cellcolor{gray!6}{0.5200} & \cellcolor{gray!6}{0.009580} & \cellcolor{gray!6}{0.2160} & \cellcolor{gray!6}{0.4652238} & \cellcolor{gray!6}{Cloudy skies} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{99.0}\\
\cmidrule{1-7}
0.1780 & 0.122000 & 0.0806 & 0.2839598 & Cloudy skies & Spring & 87.8\\
\cmidrule{1-7}
\cellcolor{gray!6}{0.1940} & \cellcolor{gray!6}{0.036800} & \cellcolor{gray!6}{0.1470} & \cellcolor{gray!6}{0.3829017} & \cellcolor{gray!6}{Cloudy skies} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{96.3}\\
\cmidrule{1-7}
0.3470 & 0.100000 & 0.0935 & 0.3057662 & Cloudy skies & Autumn & 90.0\\
\bottomrule
\end{tabu}
\end{table}
